---
title: "Hands-on-Exercise 6"
description: |
  In hands-on exercise 6, I learnt how to derive spatial weights by using functions provided by spdep package and how to apply these spatial weights to compute geographically referenced attributes.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 09-18-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1 The Study Area and Data

Two data sets will be used in this hands-on exercise, they are:

- Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.
- Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

## 1.2 Getting Started

```{r eval=TRUE, echo = TRUE}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# 2  Getting the Data into R Environment

## 2.1 Import shapefile into r environment

```{r eval=TRUE, echo = TRUE}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```

## 2.2 Import csv file into r environment

```{r eval=TRUE, echo = TRUE}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")

```

## 2.3 Performing relational join

- Used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe.
- This is performed by using left_join() of dplyr package.

```{r eval=TRUE, echo = TRUE}
hunan <- left_join(hunan,hunan2012)

```

## 2.4 Visualising Regional Development Indicator

- Prepare a choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmap package.

```{r eval=TRUE, echo = TRUE}
qtm(hunan, "GDPPC")
```

# 3 Computing Contiguity Spatial Weights

- Use poly2nb() of spdep package to compute contiguity weight matrices for the study area.
- This function builds a neighbours list based on regions with contiguous boundaries.

## 3.1 Computing (QUEEN) contiguity based neighbours

- Default for 'queen' argument is TRUE. If do not specify to queen = FALSE, this function will return a list of first order neighbours using the Queen criteria.

```{r eval=TRUE, echo = TRUE}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

- Report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours.
- There are two area units with only one neighbours

<br>

- For each polygon in our polygon object, wm_q lists all neighboring polygons.
- To see the neighbors for the first polygon in the object, type:


```{r eval=TRUE, echo = TRUE}
wm_q[[1]]

```

- Results above show that Polygon 1 has 5 neighbors.
- The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class

<br>

- To retrive the county name of Polygon ID=1:

```{r eval=TRUE, echo = TRUE}
hunan$County[1]
```

- To reveal the county names of the five neighboring polygons: 

```{r eval=TRUE, echo = TRUE}
hunan$NAME_3[c(2,3,4,57,85)]
```

- Display the complete weight matrix by using str()

```{r eval=TRUE, echo = TRUE}
str(wm_q)
```

**NOTE:** Save the trees if you are going to print out the report.

## 3.2 Computing (ROOK) contiguity based neighbours

```{r eval=TRUE, echo = TRUE}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

- The summary report above shows that there are 88 area units in Hunan.
- The most connect area unit has 10 neighbours.
- There are two area units with only one neighbours.

## 3.3 Visualising contiguity weights

- Since we are working with polygons, we need ot get points in order to make the connectivity graphs. 
- Most typical method: polygon centroids
 
    + Calculate these in the sf package before moving onto the graphs.
    + Getting Latitude and Longitude of Polygon Centroids

- Mapping function:  applies a given function to each element of a vector and returns a vector of the same length.
- Our input vector will be the geometry column of us.bound.
- Our function will be st_centroid.
- We will be using map_dbl variation of map from the purrr package.
- To get our longitude values we map the st_centroid function over the geometry column of us.bound and access the longitude value through double bracket notation [[]] and 1.
- This allows us to get only the longitude, which is the first value in each centroid.

```{r eval=TRUE, echo = TRUE}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])

```

- Do the same for latitude with one key difference.
- We access the second value per each centroid with [[2]]

```{r eval=TRUE, echo = TRUE}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])

```

- Use cbind to put longitude and latitude into the same object.

```{r eval=TRUE, echo = TRUE}
coords <- cbind(longitude, latitude)

```

- Check the first few observations to see if things are formatted correctly

```{r eval=TRUE, echo = TRUE}
head(coords)

```

### Plot Queen & Rook contiguity based neighbours maps


```{r eval=TRUE, echo = TRUE}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## 3.4 Computing distance based neighbours

- Learn how to derive distance-based weight matrices by using dnearneigh() of spdep package
- The function identifies neighbours of region points by Euclidean distance with a distance band with lower d1= and upper d2= bounds controlled by the bounds= argument.
- If unprojected coordinates are used and either specified in the coordinates object x or with x as a two column matrix and longlat=TRUE, great circle distances in km will be calculated assuming the WGS84 reference ellipsoid.

## 3.4.1 Determine the cut-off distance

- To determine the upper limit for distance band:

    + Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using knearneigh() of spdep.
    + Convert the knn object returned by knearneigh() into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using knn2nb().
    + Return the length of neighbour relationship edges by using nbdists() of spdep. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.
    + Remove the list structure of the returned object by using unlist().


```{r eval=TRUE, echo = TRUE}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

-  Report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

## 3.4.2 Computing fixed distance weight matrix

```{r eval=TRUE, echo = TRUE}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

Quiz: What is the meaning of “Average number of links: 3.681818” shown above?
- Of all regions, the average number of links for all regions are 3.681818. 

<br>

- use str() to display the content of wm_d62 weight matrix

```{r eval=TRUE, echo = TRUE}
str(wm_d62)

```

- Another way to display the structure of the weight matrix is to combine table() and card() of spdep.

```{r eval=TRUE, echo = TRUE}
table(hunan$County, card(wm_d62))

```



```{r eval=TRUE, echo = TRUE}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```


```{r eval=TRUE, echo = TRUE}
n_comp <- n.comp.nb(wm_d62)
n_comp$nc
```

## 3.4.3 Plotting fixed distance weight matrix

```{r eval=TRUE, echo = TRUE}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

- The red lines above show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

<br>

- Alternatively, we can plot both of them next to each other

```{r eval=TRUE, echo = TRUE}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

## 3.4.4 Computing adaptive distance weight matrix

- One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours.
- Having many neighbours smoothes the neighbour relationship across more neighbours.

- It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry

```{r eval=TRUE, echo = TRUE}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

- Can also display the content of the matrix by using str().

```{r eval=TRUE, echo = TRUE}
str(knn6)

```
- Notice that each county has six neighbours, no less no more

## 3.4.5 Plotting distance based neighbours

- plot the weight matrix

```{r eval=TRUE, echo = TRUE}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## 3.5 Weights based on IDW

- Learn how to derive a spatial weight matrix based on Inversed Distance method.

<br>

- Compute the distances between areas by using nbdists() of spdep.

```{r eval=TRUE, echo = TRUE}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## 3.6 Row-standardised weights matrix

- We need to assign weights to each neighboring polygon.
- In our case, each neighboring polygon will be assigned equal weight (style=“W”).
- This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values.
- While this is the most intuitive way to summaries the neighbors’ values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we’ll stick with the style=“W” option for simplicity’s sake but note that other more robust options are available, notably style=“B”

<br>

- The zero.policy=TRUE option allows for lists of non-neighbors.
- This should be used with caution since the user may not be aware of missing neighbors in their dataset however, a zero.policy of FALSE would return an error

```{r eval=TRUE, echo = TRUE}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

-  To see the weight of the first polygon’s four neighbors type

```{r eval=TRUE, echo = TRUE}
rswm_q$weights[10]
```

- Each neighbor is assigned a 0.2 of the total weight.
- This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied.

<br>

- Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.

```{r eval=TRUE, echo = TRUE}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

```{r eval=TRUE, echo = TRUE}
rswm_ids$weights[1]
```

```{r eval=TRUE, echo = TRUE}
summary(unlist(rswm_ids$weights))

```

# 4 Application of Spatial Weight Matrix

- Learn how to create four different spatial lagged variables, they are:

    + spatial lag with row-standardized weights,
    + spatial lag as a sum of neighbouring values,
    + spatial window average, and spatial window sum.

## 4.1 Spatial lag with row-standardized weights

- Finally, we’ll compute the average neighbor GDPPC value for each polygon.
- These values are often referred to as spatially lagged values


```{r eval=TRUE, echo = TRUE}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

- We can append the spatially lag GDPPC values onto hunan SpatialPolygonDataFram

```{r eval=TRUE, echo = TRUE}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

- The following table shows the average neighboring income values (stored in the Inc.lag object) for each county.

```{r eval=TRUE, echo = TRUE}
head(hunan)

```
- Next, we will plot both the GDPPC and spatial lag GDPPC for comparison

```{r eval=TRUE, echo = TRUE}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)

```

## 4.2 Spatial window sum

- The spatial window sum uses and includes the diagonal element.
- To begin, we will assign knn6 to a new variable because we will directly alter its structure to add the diagonal elements

```{r eval=TRUE, echo = TRUE}
knn6a <- knn6

```

- To add the diagonal element to the neighbour list, we just need to use include.self() from spdep.

```{r eval=TRUE, echo = TRUE}
include.self(knn6a)

```

- Next, we will assign binary weights to the neighbour structure that includes the diagonal element.

```{r eval=TRUE, echo = TRUE}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

- Again, we use nb2listw() and glist() to explicitly assign weight values.

```{r eval=TRUE, echo = TRUE}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")

```

- With our new weight structure, we can compute the lag variable with lag.listw().

```{r eval=TRUE, echo = TRUE}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)

```

- Next, we will convert the lag variable listw object into a data.frame by using as.data.frame().

```{r eval=TRUE, echo = TRUE}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
```

**Note:** The third command line on the code chunk above renames the field names of lag_knn6.res object into NAME_3 and lag_sum GDPPC respectively.

- Next, the code chunk below will be used to append lag_sum GDPPC values onto hunan sf data.frame by using left_join() of dplyr package.

```{r eval=TRUE, echo = TRUE}
hunan <- left_join(hunan, lag_knn6.res)

```

- Lastly, qtm() of tmap package is used to plot the GDPPC and lag_sum GDPPC map next to each other for quick comparison.

```{r eval=TRUE, echo = TRUE}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```
 **NOTE:** Note: For more effective comparison, it is advicible to use the core tmap mapping functions.




