---
title: "In-class Exercise 3 - 2"
description: |
  A short description of the post.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 08-31-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Load required packages
```{r echo=TRUE, eval=TRUE}
packages <- c('sf', 'tidyverse', 'readr', 'tmap')
for (p in packages){
  if (!require(p, character.only = T)) {
    install.packages(p)
  }
  library(p, character.only = T)
}
```

## 2. Import Data
```{r echo=TRUE, eval=TRUE}
mpszpop2020 <- read_rds("data/rds/mpszpop2020.rds")

```

## 3. Boxplot to visualise extreme values
```{r echo=TRUE, eval=TRUE}
ggplot(data=mpszpop2020,
        aes(x = "",
          y = AGED)) +
    geom_boxplot()
```

```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  drop_na()
```


```{r echo=TRUE, eval=TRUE}
percent <- c(0, .01, .1, .5, .9, .99, 1)
var <- mpszpop2020a["DEPENDENCY"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

## 4. Create get.var function
```{r echo=TRUE, eval=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```


## 5. Create percentile mapping function
```{r echo=TRUE, eval=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map") {
  percent <- c(0, .01, .1, .5, .9, .99, 1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(mpszpop2020) + 
  tm_polygons() +
  tm_shape(df) +
    tm_fill(vnam, 
            title =legtitle, 
            breaks = bperc,
            palette="Blues", 
            labels=c("< 1%", "1% - 10%",
                     "10% - 50%",
                     "50% - 90%",
                     "90% - 99%",
                     "> 99%")) +
    tm_borders() +
    tm_layout(title = mtitle,
              title.position = c("right", "bottom"))
}
```


### Test the percentile mapping function - 1
```{r echo=TRUE,  eval=TRUE}
percentmap("YOUNG", mpszpop2020a)
```

### Test the percentile mapping function - 2
```{r echo=TRUE,  eval=TRUE}
percentmap("AGED", mpszpop2020a)

```

*Note: Boxmap is able to show outliers. Percentile map only shows the percentage, not outliers.

## 6. Create box breaks function
```{r echo=TRUE,  eval=TRUE}
boxbreaks <- function(v, mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[4] - mult * iqr
  
  # initialize break points vector
  bb <- vector(mode="numeric", length=7)
  # logic for lower and upper fences
  if(lofence < qv[1]) { # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- celing(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```


### Test boxbreaks function 
```{r echo=TRUE,  eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  filter(AGED>=0)
var <- get.var("AGED", mpszpop2020a)
boxbreaks(var)
```
## 7. Create Boxmap function
```{r echo=TRUE,  eval=TRUE}
boxmap <- function(vnam, df,
                  legtitle=NA,
                  mtitle="Box Map",
                  mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_fill(vnam,title=legtitle,
            breaks=bb,
            palette="Blues",
          labels = c("lower outlier",
                      "< 25%",
                      "25% - 50%",
                      "50% - 75%",
                      "> 75%",
                      "upper outlier")) +
  tm_borders() +
  tm_layout(title = mtitle,
    title.position = c("right",
                        "bottom"))
}
```

## 8. Box map of AGED Population
```{r echo=TRUE,  eval=TRUE}
boxmap("ECONOMY ACTIVE", mpszpop2020a)
```


## 9. Map distribution of total population 2020
```{r echo=TRUE,  eval=TRUE}
boxmap("TOTAL", mpszpop2020a)
```

## 10. Map distribution of aged population 2020
```{r echo=TRUE,  eval=TRUE}
boxmap("AGED", mpszpop2020a)

```


## 11. Raw rate Map
### Compute raw rate
```{r echo=TRUE,  eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  mutate(`AGED%` = (`AGED`/`TOTAL`)*100) %>%
  filter(`AGED%` >= 0)
```

### Use boxmap function to plot Raw Rate Map 
```{r echo=TRUE,  eval=TRUE}
var <- get.var("AGED%", mpszpop2020a)
boxbreaks(var)
boxmap("AGED%", mpszpop2020a)
```
