---
title: "Hands-on-Exercise-5"
description: |
  A short description of the post.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 09-12-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



# 1 Installing and launching the R packages

```{r echo=TRUE, eval=TRUE}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')

for (p in packages){
if(!require(p, character.only = T)){
  install.packages(p)
}
  library(p,character.only = T)
}
```

# 2 Data Import and Preparation

- readOGD() of rgdal package to import Punggol_St and Punggol_CC geospatial data sets as SpatialLinesDataFrame and SpatialPointsDataFrame respectively

```{r echo=TRUE, eval=TRUE}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```


- Print the content of network SpatialLineDataFrame and childcare SpatialPointsDataFrame

```{r echo=TRUE, eval=TRUE}
str(network)
str(childcare)
```

## Assign CRS

- To assign EPSG code to the SpatialDataFrames. The epsg:3414 is the code for svy21.

```{r echo=TRUE, eval=TRUE}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

