---
title: "Hands-on-Exercise-5"
description: |
  In Hands-on Exercise 5, I have learnt the basic concepts of on- and off- network spatial point processes and a family of spatial statistics methods for analysing network constrained spatial point processes.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 09-12-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



# 1 Installing and launching the R packages

```{r echo=TRUE, eval=TRUE}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')

for (p in packages){
if(!require(p, character.only = T)){
  install.packages(p)
}
  library(p,character.only = T)
}
```

# 2 Data Import and Preparation

- readOGD() of rgdal package to import Punggol_St and Punggol_CC geospatial data sets as SpatialLinesDataFrame and SpatialPointsDataFrame respectively

```{r echo=TRUE, eval=TRUE}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```


- Print the content of network SpatialLineDataFrame and childcare SpatialPointsDataFrame

```{r echo=TRUE, eval=TRUE}
str(network)
str(childcare)
```

## Assign CRS

- To assign EPSG code to the SpatialDataFrames. The epsg:3414 is the code for svy21.

```{r echo=TRUE, eval=TRUE}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```


# 3 Visualising the Geospatial Data

- Good practice to visualise the geospatial data

```{r echo=TRUE, eval=TRUE}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```

- To visualise the geospatial data with high cartographic quality and interactive manner

```{r echo=TRUE, eval=TRUE}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()
```

- Set back to plot mode

```{r echo=TRUE, eval=TRUE}
tmap_mode('plot')
```


# 4 Network Constrained KDE (NetKDE) Analysis

## 4.1 Preparing the lixels objects

- Before computing NetKDE, the SpatialLines object need to be cut into lixels with a specified minimal distance.

    + The length of a lixel, lx_length is set to 700m, and
    + The minimum length of a lixel, mindist is set to 350m.
    
- After cutting, if the length of the final lixel is shorter than the minimum distance, then it is added to the previous lixel.
- If NULL, then mindist = maxdist/10. Also note that the segments that are already shorter than the minimum distance are not modified

**Note**: There is another function called lixelize_lines.mc() which provide multicore support.

```{r echo=TRUE, eval=TRUE}
lixels <- lixelize_lines(network,700,mindist = 350)
```


## 4.2 Generating line centre points

- To generate a SpatialPointsDataFrame (i.e. samples) with line centre points

```{r echo=TRUE, eval=TRUE}
samples <- lines_center(lixels)

```

- The points are located at center of the line based on the length of the line.


## 4.3 Performing NetKDE

- Compute NETKDE

Understanding the code:

- Other kernel_name available: triangle, gaussian, scaled gaussian, tricube, cosine ,triweight, epanechnikov or uniform.
- method argument:

    + simple:  The distances between events and sampling points are replaced by network distances, and the formula of the kernel is adapted to calculate the density over a linear unit instead of an areal unit.
    + discontinuous: equally “divides” the mass density of an event at intersections of lixels.(If the discontinuous method is unbiased, it leads to a discontinuous kernel function which is a bit counter-intuitive.)
    + continuous: divide the mass of the density at intersection but adjusts the density before the intersection to make the function continuous.
    
```{r echo=TRUE, eval=TRUE}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

# 5 Visualising NetKDE

## 5.1 Insert the computed density values

- Insert the computed density values (i.e. densities) into samples and lixels objects as density field.

```{r echo=TRUE, eval=TRUE}
samples$density <- densities
lixels$density <- densities
```



## 5.2 Rescale the density values

- Since svy21 projection system is in meter, the computed density values are very small i.e. 0.0000005. 
- Resale the density values from number of events per meter to number of events per kilometer.

```{r echo=TRUE, eval=TRUE}
# rescaling to help the mapping
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```


## 5.3 Prepare interactive and high cartographic quality map visualisation.

- Interactive map below  reveals road segments (darker color) with relatively higher density of childcare centres than road segments with relatively lower density of childcare centres (lighter color)

```{r echo=TRUE, eval=TRUE}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
```


```{r echo=TRUE, eval=TRUE}
tmap_mode('plot')
```


# 6 Network Constrained G- and K-Function Analysis

- To perform complete spatial randomness (CSR) test by using kfunctions() of spNetwork package. The null hypothesis is defined as:The null hypothesis is defined as:
    + **Ho: The observed spatial point events (i.e distribution of childcare centres) are uniformly distributed over a street network in Punggol Planning Area.**
- Assumption: binomial point process which implies the hypothesis that the childcare centres are randomly and independently distributed over the street network.
- If this hypothesis is rejected, we may infer that the distribution of childcare centres are spatially interacting and dependent on each other; as a result, they may form nonrandom patterns.

<br>

Understanding the code chunk:

- lines: A SpatialLinesDataFrame with the sampling points. The geometries must be a SpatialLinesDataFrame (may crash if some geometries are invalid).
- points: A SpatialPointsDataFrame representing the points on the network. These points will be snapped on the network.
- start: A double, the start value for evaluating the k and g functions.
- end: A double, the last value for evaluating the k and g functions.
- step: A double, the jump between two evaluations of the k and g function.
- width: The width of each donut for the g-function.
- nsim: An integer indicating the number of Monte Carlo simulations required. In the above example, 50 simulation was performed. Note: most of the time, more simulations are required for inference
- resolution: When simulating random points on the network, selecting a resolution will reduce greatly the calculation time. When resolution is null the random points can occur everywhere on the graph. If a value is specified, the edges are split according to this value and the random points are selected vertices on the new network.
- conf_int: A double indicating the width confidence interval (default = 0.05).

```{r echo=TRUE, eval=TRUE}
kfun_childcare <- kfunctions(network, 
                             childcare,
                             start = 0, 
                             end = 1000, 
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)
```



The output of kfunctions() is a list with the following values:

- plotkA, a ggplot2 object representing the values of the k-function
- plotgA, a ggplot2 object representing the values of the g-function
- valuesA, a DataFrame with the values used to build the plots



```{r echo=TRUE, eval=TRUE}
kfun_childcare$plotk

```

Observations:

- The blue line is the empirical network K-function of the childcare centres in Punggol planning area.
- The gray envelop represents the results of the 50 simulations in the interval 2.5% - 97.5%.
- Because the blue line between the distance of 250m-400m are below the gray area, we can infer that the childcare centres in Punggol planning area resemble regular pattern at the distance of 250m-400m.


