---
title: "Hands-on Exercise 8"
description: |
  In Hands-on Exercise 8, I have learnt how to perform geographical segmentation by using appropriate R packages.I have also been exposed to R packages for performing cluster analysis and visualising clustering results.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 10-10-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1 The Analytical Question

In geobusiness and spatial policy, it is a common practice to delineate the market or planning area into homogeneous regions by using multivariate data. In this hands-on exercise, we are interested to delineate Shan State, Myanmar into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: Radio, Television, Land line phone, Mobile phone, Computer, and Internet at home.

# 2 Dataset

Two data sets will be used in this study. They are:

- Myanmar Township Boundary Data (i.e. myanmar_township_boundaries) : This is a GIS data in ESRI shapefile format. It consists of township boundary information of Myanmar. The spatial data are captured in polygon features.
- Shan-ICT.csv: This is an extract of The 2014 Myanmar Population and Housing Census Myanmar at the township level.

Both data sets are download from Myanmar Information Management Unit (MIMU)

# 3 Install and Load Packages

- Spatial data handling

    +sf, rgdal and spdep
    
- Attribute data handling

  + tidyverse, especially readr, ggplot2 and dplyr
  
- Choropleth mapping

    + tmap
- Multivariate data visualisation and analysis

    + coorplot, ggpubr, and heatmaply
    
- Cluster analysis

    + cluster

```{r echo=TRUE, eval=TRUE}
packages = c('rgdal', 'spdep', 'tmap', 'sf', 'ggpubr', 'cluster', 'factoextra', 'NbClust', 'heatmaply', 'corrplot', 'psych', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
    }
  library(p,character.only = T)
}
```

# 4 

