---
title: "Hands-on Exercise 8"
description: |
  In Hands-on Exercise 8, I have learnt how to perform geographical segmentation by using appropriate R packages.I have also been exposed to R packages for performing cluster analysis and visualising clustering results.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 10-10-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1 The Analytical Question

In geobusiness and spatial policy, it is a common practice to delineate the market or planning area into homogeneous regions by using multivariate data. In this hands-on exercise, we are interested to delineate Shan State, Myanmar into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: Radio, Television, Land line phone, Mobile phone, Computer, and Internet at home.

# 2 Dataset

Two data sets will be used in this study. They are:

- Myanmar Township Boundary Data (i.e. myanmar_township_boundaries) : This is a GIS data in ESRI shapefile format. It consists of township boundary information of Myanmar. The spatial data are captured in polygon features.
- Shan-ICT.csv: This is an extract of The 2014 Myanmar Population and Housing Census Myanmar at the township level.

Both data sets are download from Myanmar Information Management Unit (MIMU)

# 3 Install and Load Packages

- Spatial data handling

    + sf, rgdal and spdep
    
- Attribute data handling

    + tidyverse, especially readr, ggplot2 and dplyr
  
- Choropleth mapping

    + tmap
- Multivariate data visualisation and analysis

    + coorplot, ggpubr, and heatmaply
    
- Cluster analysis

    + cluster

```{r echo=TRUE, eval=TRUE}
packages = c('rgdal', 'spdep', 'tmap', 'sf', 'ggpubr', 'cluster', 'factoextra', 'NbClust', 'heatmaply', 'corrplot', 'psych', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
    }
  library(p,character.only = T)
}
```

# 4 Data Import and Preparation

## 4.1 Importing geospatial data into R environment

- The Myanmar Township Boundary GIS data is in ESRI shapefile format.
- Use *st_read()* function of **sf** package.
- Filter extracts particular regions in Shan

```{r echo=TRUE, eval=TRUE}
shan_sf <- st_read(dsn = "data/geospatial", layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)"))
```

## 4.2 View newly created sf dataframe

- View shan_sf 
- Since shan_sf is conformed to tidy framework, we can also *glimpse()* to reveal the data type of it’s fields.

```{r echo=TRUE, eval=TRUE}
shan_sf
glimpse(shan_sf)
```

## 4.3 Importing aspatial data into R environment

- Use *read_csv()* function of **readr** package
- It is saved in R’s * tibble data.frame* format.
- Use *summary()* to reveal the summary statistics of ict data.frame

```{r echo=TRUE, eval=TRUE}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")

summary(ict)
```

Results above show that:

- There are a total of eleven fields and 55 observation in the tibble data.frame.

## 4.4 Derive new variables using dplyr package

- The unit of measurement of the values are number of household.
- Using these values directly will be **bias** by the underlying total number of households.
- In general, the townships with relatively higher total number of households will also have higher number of households owning radio, TV, etc.
- In order to overcome this problem, we will **derive the penetration rate** of each ICT variable
- - Use *summary()* to reveal the summary statistics of ict_derived data.frame

```{r echo=TRUE, eval=TRUE}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 

summary(ict_derived)

```

# 5 Exploratory Data Analysis (EDA)

## 5.1 EDA using statistical graphics

### 5.1.1 Histograms

- Plot the distribution of the variables (i.e. Number of households with radio) 
- Histogram is useful to identify the overall distribution of the data values (i.e. left skew, right skew or normal distribution)

```{r echo=TRUE, eval=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

### 5.1.2 Boxplot

- Boxplots are useful to detect if there are outliers.

```{r echo=TRUE, eval=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_boxplot(color="black", fill="light blue")
```

### 5.1.3 Histograms (newly derived variables)

- Plot the distribution of the newly derived variables (i.e. Radio penetration rate)

```{r echo=TRUE, eval=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

### 5.1.4 Boxplot (newly derived variables)


```{r echo=TRUE, eval=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", fill="light blue")
```

### 5.1.4 Multiple Histograms (selected variables)

- Multiple histograms are plotted to reveal the distribution of the selected variables in the ict_derived data.frame
- *ggarange()* function of **ggpubr** package is used to group these histograms together.

```{r echo=TRUE, eval=TRUE}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```

## 5.2 EDA using choropleth map

### 5.2.1 Joining geospatial data with aspatial data

- Combine both the geospatial data object (i.e. shan_sf) and aspatial data.frame object (i.e. ict_derived) into one
- Use the *left_join* function of **dplyr** package
- *shan_sf* simple feature data.frame will be used as the base data object and the *ict_derived* data.frame will be used as the join table.
- The unique identifier used to join both data objects is *TS_PCODE*

```{r echo=TRUE, eval=TRUE}
shan_sf <- left_join(shan_sf, ict_derived, by=c("TS_PCODE"="TS_PCODE"))

```

- It is important to note that there is no new output data been created. Instead, the data fields from ict_derived data frame are now updated into the data frame of shan_sf.

### 5.2.2 Preparing a choropleth map

- using the *qtm()* function of **tmap** package to  look at the distribution of Radio penetration rate of Shan State at township

```{r echo=TRUE, eval=TRUE}
qtm(shan_sf, "RADIO_PR")

```

### 5.2.3 Reveal distribution (bias)

- Create two choropleth maps, one for the total number of households (i.e. TT_HOUSEHOLDS.map) and one for the total number of household with Radio (RADIO.map) to  **reveal the distribution shown in the choropleth map above are bias** to the underlying total number of households at the townships

```{r echo=TRUE, eval=TRUE}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Number Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```

Results above show that:
 
-  Townships with relatively larger number of households are also showing relatively higher number of radio ownership.

### 5.2.4 Distribution of total number of households and Radio penetration rate

```{r echo=TRUE, eval=TRUE}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```

Question: Can you identify the differences?

## 5.3 Correlation Analysis

Before we perform cluster analysis, it is important for us to ensure that the cluster variables are not highly correlated.

- Use *corrplot.mixed()* function of **corrplot** package to visualise and analyse the correlation of the input variables.The upper half shows numbers which represents the correlation.
- Darker colours means it is highly correlated.

```{r echo=TRUE, eval=TRUE}
cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

Results above show that:

- Correlation plot shows that COMPUTER_PR and INTERNET_PR are highly correlated.
- This suggest that only one of them should be used in the cluster analysis instead of both.

# 6 Hierarchy Cluster Analysis

## 6.1 Prepare variables

### 6.1.1 Extrating clustering variables

- To extract the clustering variables from the shan_sf simple feature object into data.frame.
- *st_set_geometry(NULL)* drops geometric information

```{r echo=TRUE, eval=TRUE}
cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

**Note:** Final clustering variables list **does not include variable INTERNET_PR** because it is highly correlated with variable COMPUTER_PR.

### 6.1.2 Change row labels

- Change the row label using township name instead of row number 

```{r echo=TRUE, eval=TRUE}
row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)
```

### 6.1.3 Delete TS.x field

```{r echo=TRUE, eval=TRUE}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)
```

## 6.2 Data Standardisation

In general, multiple variables will be used in cluster analysis. It is not unusual their values range are different. In order to avoid the cluster analysis result is biased to clustering variables with large values, it is useful to standardise the input variables before performing cluster analysis.

### 6.2.1 Min-Max standardisation

- *normalize()* of **heatmaply** package is used to stadardisation the clustering variables by using Min-Max method.
- *summary()* is then used to display the summary statistics of the standardised clustering variables.

```{r echo=TRUE, eval=TRUE}
shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)
```

Results above show that:

- Values range of the Min-max standardised clustering variables are **0-1** now.

### 6.2.2 Z-score standardisation

- Use *scale()* of **BaseR**
- *describe()* of **psych** package is used here instead of summary() of Base R because the earlier provides standard deviation.

```{r echo=TRUE, eval=TRUE}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)
```

Results above show that:

- mean and standard deviation of the Z-score standardised clustering variables are 0 and 1 respectively.

**Warning: Z-score standardisation method should only be used if we would assume all variables come from some normal distribution.**

### 6.2.3 Visualising the standardised clustering variables

- It is a good practice to visualise their distribution graphically.
- Plot the scaled Radio_PR field

```{r echo=TRUE, eval=TRUE, fig.width=8}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

**Note:** Notice that the overall distribution of the clustering variables will change after the data standardisation. Hence, it is advisible **NOT** to perform data standardisation if the values range of the clustering variables are not very large.

## 6.3 Computing proximity matrix

- Use *dist()* of R to calculate distance matrix
- *dist()* supports six distance proximity calculations, they are: euclidean, maximum, manhattan, canberra, binary and minkowski. The default is euclidean proximity matrix.
-  List the content of proxmat for visual inspection

```{r echo=TRUE, eval=TRUE}
proxmat <- dist(shan_ict, method = 'euclidean')
proxmat
```

## 6.4 Hierarchical clustering

### 6.4.1 Compute hierarchical clustering

- Use *hclust()* of **R** stats 
- *hclust()* employed agglomeration method to compute the cluster. Eight clustering algorithms are supported, they are: ward.D, ward.D2, single, complete, average(UPGMA), mcquitty(WPGMA), median(WPGMC) and centroid(UPGMC).
- The hierarchical clustering output is stored in an object of class hclust which describes the tree produced by the clustering process.

```{r echo=TRUE, eval=TRUE}
hclust_ward <- hclust(proxmat, method = 'ward.D')

```

### 6.4.2 Plot tree

- Use *plot()* of **R** Graphics 

```{r echo=TRUE, eval=TRUE}
plot(hclust_ward, cex = 0.6)

```

## 6.5 Select the optimal clustering algorithm

One of the challenge in performing hierarchical clustering is to identify stronger clustering structures. The issue can be solved by using use agnes() function of cluster package. It functions like hclus(), however, with the agnes() function you can also get the agglomerative coefficient, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).

- To compute the agglomerative coefficients of all hierarchical clustering algorithms.

```{r echo=TRUE, eval=TRUE}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)
```

Results above show that: 

- Ward’s method provides the strongest clustering structure among the four methods assessed.
- Hence, in the subsequent analysis, only Ward’s method will be used.


## 6.6 Determining Optimal Clusters

There are three commonly used methods to determine the optimal clusters, they are:

- Elbow Method
- Average Silhouette Method
- Gap Statistic Method

### 6.6.1 Gap Statistic Method

The gap statistic compares the total within intra-cluster variation for different values of k with their expected values under null reference distribution of the data. The estimate of the optimal clusters will be value that maximize the gap statistic (i.e., that yields the largest gap statistic). This means that the clustering structure is far away from the random uniform distribution of points.

#### 6.6.1.1 Compute gap statistic 

- Use *clusGap()* of **cluster** package
- *hcut* function used is from **factoextra** package.

```{r echo=TRUE, eval=TRUE}
set.seed(12345)
gap_stat <- clusGap(shan_ict, FUN = hcut, nstart = 25, K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
```

#### 6.6.1.2 Visualise plot

-  *fviz_gap_stat()* of **factoextra** package

```{r echo=TRUE, eval=TRUE}
fviz_gap_stat(gap_stat)

```

Results above show that:

- The recommended number of cluster to retain is 1.
- However, it is not logical to retain only one cluster. By examine the gap statistic graph, the **6-cluster** gives the largest gap statistic and should be the next best cluster to pick.

**Note:** In addition to these commonly used approaches, the NbClust package, published by Charrad et al., 2014, provides 30 indices for determining the relevant number of clusters and proposes to users the best clustering scheme from the different results obtained by varying all combinations of number of clusters, distance measures, and clustering methods.

## 6.7 Interpreting the dendrograms

In the dendrogram displayed above, each leaf corresponds to one observation. As we **move up** the tree, observations that are **similar to each other** are **combined into branches**, which are themselves fused at a higher height.

The height of the fusion, provided on the vertical axis, indicates the (dis)similarity between two observations. The **higher the height of the fusion**, the **less similar** the observations are.

Note: conclusions about the **proximity** of two observations can be 8*drawn only based on the height where branches containing those two observations first are fused**. We cannot use the proximity of two observations along the horizontal axis as a criteria of their similarity.

- Use *rect.hclust()* of **R** stats to draw the dendrogram with a border around the selected clusters.
- The argument border is used to specify the border colors for the rectangles. border is to have random colours for 2 to 5 of the clusters
- There are 6 clusters, hence k = 6

```{r echo=TRUE, eval=TRUE}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, k = 6, border = 2:5)
```

## 6.8 Visually-driven hierarchical clustering analysis

With **heatmaply** package, we are able to build both highly interactive cluster heatmap or static cluster heatmap.

### 6.8.1 Transforming the data frame into a matrix

- The data was loaded into a data frame, but it has to be a data matrix to make your heatmap.
- Use *data.matrix()*

```{r echo=TRUE, eval=TRUE}
shan_ict_mat <- data.matrix(shan_ict)

```

### 6.8.2 Plotting interactive cluster heatmap using heatmaply()

- use *heatmaply()* to build an interactive cluster heatmap.
- seriate argument is for standardization method

```{r echo=TRUE, eval=TRUE}
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )
```

## 6.9 Mapping the clusters formed

With closed examination of the dendragram above, we have decided to retain five clusters.

### 6.9.1 Derive 6-cluster model

- Use *cutree()* of **R Base**
- *as.factor()* to convert the numbers into factors if not, we will not be able to map it as a categorical data. We want to plot it as a area-shading map, instead of a choropleth map. 

```{r echo=TRUE, eval=TRUE}
groups <- as.factor(cutree(hclust_ward, k=6))

```

### 6.9.2 Append *groups*

In order to visualise the clusters, the groups object need to be appended onto shan_sf simple feature object.

- the groups list object will be converted into a matrix;
- cbind() is used to append groups matrix onto shan_sf to produce an output simple feature object called shan_sf_cluster; and
- rename of dplyr package is used to rename as.matrix.groups field to CLUSTER.

```{r echo=TRUE, eval=TRUE}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)
```

### 6.9.3 Plot choropleth map

-  *qtm()* of **tmap** package 

```{r echo=TRUE, eval=TRUE}
qtm(shan_sf_cluster, "CLUSTER")
```

Results above show that: 

- Legend is 1,2,3 as it is a categorical 
- Clusters are very fragmented.
- This is one of the major limitations when non-spatial clustering algorithm such as hierarchical cluster analysis method is used.

# 7 Spatially Constrained Clustering - SKATER approach

To derive spatially constrained cluster by using SKATER method.

## 7.1 Convert into SpatialPolygonsDataFrame

- First, we need to convert shan_sf into SpatialPolygonDataFrame. This is because SKATER function only support sp objects such as SpatialPolygonDataFrame.
- Use *as_Spatial()* of **sf** package to convert shan_sf into a SpatialPolygonDataFrame called shan_sp.

```{r echo=TRUE, eval=TRUE}
shan_sp <- as_Spatial(shan_sf)

```

## 7.2 Compute Neighbour List

- *poly2nd()* of **spdep** package will be used to compute the neighbours list from polygon list.

```{r echo=TRUE, eval=TRUE}
shan.nb <- poly2nb(shan_sp)
summary(shan.nb)
```

- Plot the neighbours list on shan_sp 
- Since we now can plot the community area boundaries as well, we plot this graph on top of the map

<br>

- Plot command gives the boundaries.
- The plot of the neighbor list object, with coordinates applied to the original SpatialPolygonDataFrame (Shan state township boundaries) to extract the centroids of the polygons.
- These are used as the nodes for the graph representation. We also set the color to blue and specify add=TRUE to plot the network on top of the boundaries.

```{r echo=TRUE, eval=TRUE}
plot(shan_sp, border=grey(.5))
plot(shan.nb, coordinates(shan_sp), col="blue", add=TRUE)
```

## 7.3 Computing minimum spanning tree

### 7.3.1 Calculating edge costs

- *nbcosts()* of **spdep** package is used to compute the cost of each edge.
- It is the distance between it nodes.
- This function compute this distance using a data.frame with observations vector in each node.

```{r echo=TRUE, eval=TRUE}
lcosts <- nbcosts(shan.nb, shan_ict)

```

For each observation, this gives the pairwise dissimilarity between its values on the five variables and the values for the neighbouring observation (from the neighbour list). Basically, this is the notion of a generalised weight for a spatial weights matrix.


Next, incorporate these costs into a weights object in the same way as we did in the calculation of inverse of distance weights. In other words, we convert the neighbour list to a list weights object by specifying the just computed lcosts as the weights.

- Use *nb2listw(*) of **spdep** package
- we specify the style as B to make sure the cost values are not row-standardised.

```{r echo=TRUE, eval=TRUE}
shan.w <- nb2listw(shan.nb, lcosts, style="B")
summary(shan.w)
```

### 7.3.2 Computing minimum spanning tree

- Minimum spanning tree is computed by mean of the *mstree()* of **spdep** package

```{r echo=TRUE, eval=TRUE}
shan.mst <- mstree(shan.w)

```

- After computing the MST, check its class and dimension by using the code chunk below.

```{r echo=TRUE, eval=TRUE}
class(shan.mst)
dim(shan.mst)

```

Results above show that:

- The dimension is 54 and not 55. This is because the minimum spanning tree consists on n-1 edges (links) in order to traverse all the nodes.

<br>

- Display the content of *shan.mst* by using *head()* 

```{r echo=TRUE, eval=TRUE}
head(shan.mst)

```

- Plot method for the MST include a way to show the observation numbers of the nodes in addition to the edge.
- As before, we plot this together with the township boundaries. We can see how the initial neighbour list is simplified to just one edge connecting each of the nodes, while passing through all the nodes.

```{r echo=TRUE, eval=TRUE}
plot(shan_sp, border=gray(.5))
plot.mst(shan.mst, coordinates(shan_sp), 
     col="blue", cex.lab=0.7, cex.circles=0.005, add=TRUE)
```

## 7.4 Computing spatially constrained clusters using SKATER method

- use *skater()* of **spdep** package to compute the spatially constrained cluster
- *skater()* takes in 3 mandatory arguments:

    + the first two columns of the MST matrix (i.e. not the cost)
    + the data matrix (to update the costs as units are being grouped)
    + the number of cuts. **Note:** It is set to one less than the number of clusters. So, the value specified is not the number of clusters, but the number of cuts in the graph, one less than the number of clusters.
    
```{r echo=TRUE, eval=TRUE}
clust6 <- skater(shan.mst[,1:2], shan_ict, method = "euclidean", 5)
```

- Examine result of the *skater()* is an object of class skater contents

```{r echo=TRUE, eval=TRUE}
str(clust6)

```

The most interesting component of this list structure is the groups vector containing the labels of the cluster to which each observation belongs (as before, the label itself is arbitary). This is followed by a detailed summary for each of the clusters in the edges.groups list. Sum of squares measures are given as ssto for the total and ssw to show the effect of each of the cuts on the overall criterion.

- Check the cluster assignment

```{r echo=TRUE, eval=TRUE}
ccs6 <- clust6$groups
ccs6
```

We can find out how many observations are in each cluster by means of the table command. Parenthetially, we can also find this as the dimension of each vector in the lists contained in edges.groups. For example, the first list has node with dimension 12, which is also the number of observations in the first cluster.

```{r echo=TRUE, eval=TRUE}
table(ccs6)

```

Lastly, we can also plot the pruned tree that shows the five clusters on top of the townshop area.

```{r echo=TRUE, eval=TRUE}
plot(shan_sp, border=gray(.5))
plot(clust6, coordinates(shan_sp), cex.lab=.7,
     groups.colors=c("red","green","blue", "brown", "pink"), cex.circles=0.005, add=TRUE)
```

## 7.5 Visualising the clusters in choropleth map

- To plot the newly derived clusters by using SKATER method.

```{r echo=TRUE, eval=TRUE}
groups_mat <- as.matrix(clust6$groups)
shan_sf_spatialcluster <- cbind(shan_sf_cluster, as.factor(groups_mat)) %>%
  rename(`SP_CLUSTER`=`as.factor.groups_mat.`)
qtm(shan_sf_spatialcluster, "SP_CLUSTER")
```

For easy comparison, it will be better to place both the hierarchical clustering and spatially constrained hierarchical clustering maps next to each other.

```{r echo=TRUE, eval=TRUE}
hclust.map <- qtm(shan_sf_cluster,
                  "CLUSTER") + 
  tm_borders(alpha = 0.5) 

shclust.map <- qtm(shan_sf_spatialcluster,
                   "SP_CLUSTER") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(hclust.map, shclust.map,
             asp=NA, ncol=2)
```



