---
title: "Hands-on Exercise 11"
description: |
  In Hands-on Exercise 11, I have learnt how to model geographical accessibility using R Packages.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 10-30-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1 Introduction

In this hands-on exercise, I have gain hands-on experience on how to model geographical accessibility by using R’s geospatial analysis packages.

# 2 Data

Four data sets will be used in this hands-on exercise, they are:

- MP14_SUBZONE_NO_SEA_PL: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

- hexagons: A 250m radius hexagons GIS data. This data set was created by using st_make_grid() of sf package. It is in ESRI shapefile format.

- ELDERCARE: GIS data showing location of eldercare service. This data is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

- OD_Matrix: a distance matrix in csv format. There are six fields in the data file. They are:

    + origin_id: the unique id values of the origin (i.e. fid of hexagon data set.),
    + destination_id: the unique id values of the destination (i.e. fid of ELDERCARE data set.),
    + entry_cost: the perpendicular distance between the origins and the nearest road),
    + network_cost: the actual network distance from the origin and destination,
    + exit_cost: the perpendicular distance between the destination and the nearest road), and
    + total_cost: the summation of entry_cost, network_cost and exit_cost.
All the values of the cost related fields are in metres.

# 3 Install and Launch R Packages

The R packages need for this exercise are as follows:

- Spatial data handling

    + sf

- Modelling geographical accessibility

    + spatialAcc

- Attribute data handling

    + tidyverse, especially readr and dplyr

- thematic mapping

    + tmap

- Staistical graphic

    + ggplot2

- Statistical analysis

    + ggstatsplot

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
packages = c('tmap', 'SpatialAcc',
             'sf', 'ggstatsplot', 'reshape2',
             'tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

Notice that with tidyverse, it also installs other R packages such as readr, dplyr and ggplots, tidyr, stringr, forcats, tibble, purrr and magrittr packages.

# 4 Geospatial Data Wrangling

## 4.1 Import geospatial data

- *st_read()* of **sf** package to import these three data sets 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 

```

## 4.2 Update CRS information

### 4.2.1 Assign correct code

- *st_transform()* updates the newly imported mpsz with the correct ESPG code (i.e. 3414)

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

### 4.2.2 Verify CRS

- *st_crs()* of **sf**  package to verify the projection of the newly transformed mpsz_svy21

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
st_crs(mpsz)

```

## 4.3 Clean and update attribute fields of the geospatial data

- Exclude redundant fields
- *mutate()*  of **dplyr** package to create  new field called demand and capacity that will be added into the data table of hexagons and eldercare sf data frame respectively. 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)

hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

# 5 Apsaital Data Handling and Wrangling

## 5.1 Import Distance Matrix

-  *read_cvs()* of **readr** package to import OD_Matrix.csv.
- The imported object is a tibble data.frame called ODMatrix.
- Note: skip = 0 is not exactly important here since we have already prepared the data such that there are rows and columns

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

## 5.2 Tidy distance matrix

### 5.2.1 Transform matrix

- The imported ODMatrix organised the distance matrix columnwise.
- On the other hand, most of the modelling packages in R is expecting where rows represent origins (i.e. also know as from field) and the columns represent destination (i.e. also known as to field.)
- *spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.
- **Note**: We can also use *pivot_wider()* instead of *spread()*

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

### 5.2.2 Convert from m to km

- *as.matrix()* of **Base R** to convert unit of measurement from m to km

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
distmat_km<-as.matrix(distmat/1000)
```

# 6 Modelling and Visualising Accessibility using Hansen Method

## 6.1 Computing Hansen’s accessibility

- *ac()* of **SpatialAcc** package to compute Hansen’s accessibility
- *data.frame()* is used to save the output in a data frame called acc_Handsen.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
acc_Hansen <- data.frame(ac(hexagons$demand, # hexcen$demand
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

## 6.2 Rename column name

- *colnames()* of **Base R** to rename column names

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
colnames(acc_Hansen) <- "accHansen"

```

## 6.3 Convert to tibble format

- *tbl_df()* of **dplyr** package to convert data table into tibble format

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
acc_Hansen <- tbl_df(acc_Hansen)

```

## 6.4 Join *acc_Hansen* with *hexagons* 

- *bind_cols()* of **dplyr** will be used to join the acc_Hansen tibble data frame with the hexagons simple feature data frame.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)

```

## 6.5 Compiled version of the above steps

```{r echo=TRUE, eval=FALSE,  cache=TRUE}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- tbl_df(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)

```

## 6.6 Visualising Hansen’s accessibility

### 6.6.1 Extracting map extend

- *st_bbox()* of **sf** package to extract the extend of hexagons simple feature data frame

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mapex <- st_bbox(hexagons) #view extent

```

### 6.6.2 Plot eldercare

- Uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## 6.7 Statistical graphic visualisation
In this section, we are going to compare the distribution of Hansen’s accessibility values by URA Planning Region.

### 6.7.1 Add the planning region field

- *st_join()* of **sf** package to add the planning region field into *haxegon_Hansen* simple feature data frame

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

### 6.7.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# 7 Modelling and Visualising Accessibility using KD2SFCA Method

## 7.1 Computing KD2SFCA’s accessibility

- The code chunk below calculates Hansen’s accessibility using ac() of SpatialAcc and data.frame() is used to save the output in a data frame called acc_KD2SFCA.
- Notice that KD2SFCA is used for family argument.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))

colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- tbl_df(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## 7.2 Visualising KD2SFCA’s accessibility

- The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.
Notice that mapex is reused for bbox argument.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```


## 7.2 Statistical graphic visualisation
Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

### 7.2.1 Add the planning region field

- - *st_join()* of **sf** package to add the planning region field into *hexagon_KD2SFCA* simple feature data frame

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

### 7.2.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```
# 8 Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method

## 8.1 Computing SAM accessibility

- *ac()* of **SpatialAcc** and *data.frame()* is used to save the output in a data frame called acc_SAM.
- Notice that SAM is used for family argument.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- tbl_df(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## 8.2 Visualising SAM’s accessibility

- The code chunk below uses a collection of mapping fucnfunctionstions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.
- Notice that mapex is reused for bbox argument.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## 8.3 Statistical graphic visualisation
Now, we are going to compare the distribution of SAM accessibility values by URA Planning Region.

### 8.3.1 Add the planning region field

- *st_join()* of **sf** package to add the planning region field into *hexagon_SAM * simple feature data frame

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

### 8.3.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```







