---
title: "Hands-on Exercise 11"
description: |
  In Hands-on Exercise 11, I have learnt how to model geographical accessibility using R Packages.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 10-30-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1 Introduction

In this hands-on exercise, I have gain hands-on experience on how to model geographical accessibility by using R’s geospatial analysis packages.

# 2 Data

Four data sets will be used in this hands-on exercise, they are:

- MP14_SUBZONE_NO_SEA_PL: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.

- hexagons: A 250m radius hexagons GIS data. This data set was created by using st_make_grid() of sf package. It is in ESRI shapefile format.

- ELDERCARE: GIS data showing location of eldercare service. This data is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.

- OD_Matrix: a distance matrix in csv format. There are six fields in the data file. They are:

    + origin_id: the unique id values of the origin (i.e. fid of hexagon data set.),
    + destination_id: the unique id values of the destination (i.e. fid of ELDERCARE data set.),
    + entry_cost: the perpendicular distance between the origins and the nearest road),
    + network_cost: the actual network distance from the origin and destination,
    + exit_cost: the perpendicular distance between the destination and the nearest road), and
    + total_cost: the summation of entry_cost, network_cost and exit_cost.
All the values of the cost related fields are in metres.

# 3 Install and Launch R Packages

The R packages need for this exercise are as follows:

- Spatial data handling

    + sf

- Modelling geographical accessibility

    + spatialAcc

- Attribute data handling

    + tidyverse, especially readr and dplyr

- thematic mapping

    + tmap

- Staistical graphic

    + ggplot2

- Statistical analysis

    + ggstatsplot

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
packages = c('tmap', 'SpatialAcc',
             'sf', 'ggstatsplot', 'reshape2',
             'tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

Notice that with tidyverse, it also installs other R packages such as readr, dplyr and ggplots, tidyr, stringr, forcats, tibble, purrr and magrittr packages.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```

```{r echo=TRUE, eval=TRUE,  cache=TRUE}

```







