---
title: "Hands-on Exercise 10"
description: |
  In Hands-on Exercise 10, I have learnt how to calibrate spatial interaction models by using GLM() of Base R.(contains additional notes taken in class)
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 10-23-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1 Overview

In this hands-on exercise, I have gained hands-on experience on how to calibrate Spatial Interaction Models (SIM) by using GLM() of Base R. The use case is adapted from Modelling population flows using spatial interaction models by Adam Dennett.

# 2 Data

Two data sets will be used in this hands-on exercise, they are:

- [Greater Capital City Statistical Areas](https://www.abs.gov.au/websitedbs/censushome.nsf/home/factsheetsgeography/$file/Greater%20Capital%20City%20Statistical%20Area%20-%20Fact%20Sheet.pdf), Australia. It is in geojson format.
- [Migration data from 2011 Australia Census](https://www.abs.gov.au/ausstats/abs@.nsf/ViewContent?readform&view=productsbytopic&Action=Expand&Num=5.5.5). It is in csv file format.

# 3 Install and Launch R Packages

The R packages need for this exercise are as follows:

- Spatial data handling

    + sf, sp, ‘geojsonio’, ‘stplanr’

- Attribute data handling

    + tidyverse, especially readr and dplyr, reshape2,

- Thematic mapping

    + tmap

- Statistical graphic

    + ggplot2

- Statistical analysis

    + caret (can also be used for data mining techniques)
    
```{r echo=TRUE, eval=TRUE,  cache=TRUE}
packages = c('tmap', 'tidyverse',
             'sp', 'caret',
             'geojsonio', 'stplanr',
             'reshape2', 'broom', 'sf')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

```{r echo=TRUE, eval=FALSE,  cache=TRUE}
library(devtools)
install_version("sf", version = "0.9-8", repos = "http://cran.us.r-project.org")
```

Note: stplanr was removed from cran recently. The latest version is 0.8.4. 

```{r echo=TRUE, eval=FALSE,  cache=TRUE}
library(sf)
library(stplanr)
```



# 4 Geospatial Data

## 4.1 Download Data
- Download a copy of Greater Capital City Statistical Areas boundary layer from a dropbox depository by using geojson_read() of geojsonio package

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")

```

## 4.2 Extract Data

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
Ausdata <- Aus@data
```

## 4.3 Convert to sf object and set CRS

- Use *st_as_sf()* of **sf** package to convert to sf object
- Use *st_set_crs()* of **sf** package to set CRS

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
```

## 4.4 Check simple features

- Use *st_is_valid()* of **sf** package to check if all the simple features are valid 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
st_is_valid(AusSF)
```

Results above show that:

- There are several invalid features

## 4.5 Fix simple features

- Use *st_make_valid()* of **sf** package to fix features

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
AusSF <- st_make_valid(AusSF)
```

## 4.6 Check simple features again

- Use *st_is_valid()* of **sf** package to ensure that all the simple features are valid 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
st_is_valid(AusSF)
```

## 4.7 Display boundary layer

- Use  *qtm()* of **tmap** package to plot AusSF simple feature data.frame

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
tmap_mode("plot")
qtm(AusSF)
```

## 4.8 Display Data Table

- Use head to view the first few rows of data table

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
head(AusSF, 10)
```

Results above show that:

- With close examination, the code is not in order.
- We have to reorder the code column

## 4.9 Order data table

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
AusSF1 <- AusSF[order(AusSF$GCCSA_CODE),]
head(AusSF1, 10)
```

## 4.10 Convert to sp object

- Use *as()* with argument **'Spatial'** to convert to sp object

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
Aus <- as(AusSF1, "Spatial")

```

## 4.11 Calculate Distance Matrix

In our spatial interaction model, space is one of the key predictor variables. In this example we will use a very simple Euclidean distance measure between the centroids of the Greater Capital City Statistical Areas as our measure of space.

Caution note: With some areas being huge, there are obvious potential issues with this (for example we could use the average distance to larger settlements in the noncity areas). However as this is just an example and will proceed with a simple solution for now.

### 4.11.1 Re-project to projected coordinate system

- The original data is in geographical coordinate system and the unit of measurement is in decimal degree, which is not appropriate for distance measurement. Before we compute the distance matrix, we will re-project the Aus into projected coordinate system.
- Use *spTransform()* of **sp** package

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
AusProj <- spTransform(Aus,"+init=epsg:3112")
summary(AusProj)
```

### 4.11.2 Compute distance matrix

- Technically, we can use *st_distance()* of **sf** package to compute the distance matrix.(have to use *st_centroid()* to calculate its centroid first)
- However, since the process took much longer time to complete, *spDist()* of **sp** package is used instead. 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
dist <- spDists(AusProj)
dist 
```

### 4.11.3 Convert distance matrix into distance pair list

- In order to integrate the distance matrix with the migration flow data.frame (in the later steps), we need to transform the newly derived distance matrix into a three columns distance values list.
- Use *melt()* of **reshape2** package of R to complete the task (we can also achieve the same task by using *pivot_longer()* of **dplyr** package).

Note: **reshape2** is an old package. Should use **tidyr** or **dplyr** when we can.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
distPair <- melt(dist)
head(distPair, 10)
```

### 4.11.4 Convert unit of measurement from metres into km

- The unit of measurement of Australia projected coordinate system is in metre. As a result, the values in the distance matrix are in metres too. -  Convert the distance values into kilometres.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

# 5 Interaction Data

## 5.1 Import interaction data

- Use *read_csv()* of **readr** package
- Note: if there are any NA values, do not remove it. Instead, replace it with 0 to retain the flow. 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)
glimpse(mdata)
```


## 5.2 Combine the imported migration data

We need to add in our distance data that we generated earlier and create a new column of total flows which excludes flows that occur within areas (we could keep the within-area (intra-area) flows in, but they can cause problems so we will just exclude them  for now ).

### 5.2.1 Create new column and set to small number

- Create a new total column which excludes intra-zone flow totals. Set them to a very very small number to avoid making the intra-zonal distance becoming 0.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
```

### 5.2.2 Join with flow data

- We ordered our spatial data earlier so that our zones are in their code order. We can now easily join these data together with our flow data as they are in the correct order.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdata$dist <- distPair$value 

```

### 5.2.3 Set intra-zonal distances to 0

-  While we are here, rather than setting the intra-zonal distances to 0, we should set them to something small (most intrazonal moves won’t occur over 0 distance)
- Use *glimpse()* to look at the new data

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
glimpse(mdata)
```

# 6 Visualise with Desire Line

In this section, you will learn how to prepare a desire line by using **stplanr** package.

## 6.1 Remove intra-zonal flows

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]

```

## 6.2 Remove columns

- Use the *od2line()* function **stplanr** package to remove all but the origin, destination and flow columns.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, zones = Aus)
```

## 6.3 Convert  flows to WGS84 projection

- Use *spTransform()* of **sp** package

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326" )

```

## 6.4 Convert Aug later to WGS84 projection

- - Use *spTransform()* of **sp** package

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
AusWGS <- spTransform(Aus,"+init=epsg:4326" )

```

## 6.5 Set line widths

- Set the line widths to some sensible value according to the flow

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10

```

## 6.6 Plot desire line map

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
plot(travel_networkwgs, lwd = w)
plot(AusWGS, add=T)
```



# 7 Build Spatial Interaction Models

Learn how to use R Stat function to calibrate the Spatial Interaction Models. Instead of using *lm()*, *glm()* **function will be used**. This is because *glm()* allows us to **calibrate the model using generalised linear regression methods**.

## 7.1 Unconstrained Spatial Interaction Model

In this section, we will calibrate an unconstrained spatial interaction model by using glm(). The explanatory variables are origin population (i.e. vi1_origpop), destination median income (i.e. wj3_destmedinc) and distance between origin and destination in km (i.e. dist).

### 7.1.1 Caliberate model

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(uncosim)
```

Result above show that:

- The parameter estimates of the explanatory variables are significant at alpha value 0.001.

### 7.1.2 Fit the model

- Use *fitted()* of **R** to compute the fitted values to assess the performance of the model

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$fitted <- fitted(uncosim)

```

### 7.1.3 A more difficult way (optional)

Another way to calculate the estimates is to plug all of the parameters back into Equation 6

- Assign the parameter values from the model to the appropriate variables

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
k <- uncosim$coefficients[1]
mu <- uncosim$coefficients[2]
alpha <- uncosim$coefficients[3]
beta <- -uncosim$coefficients[4]

```

- Next, plug everything back into the Equation 6 model (be careful with the positive and negative signing of the parameters as the beta parameter may not have been saved as negative so will need to force negative)
- Below are 2 ways to do it (1 of them are commented out)

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$unconstrainedEst2 <- exp(k+(mu*log(mdatasub$vi1_origpop))+(alpha*log(mdatasub$wj3_destmedinc))-(beta*log(mdatasub$dist)))

# mdatasub$unconstrainedEst2 <- (exp(k)*exp(mu*log(mdatasub$vi1_origpop))*exp(alpha*log(mdatasub$wj3_destmedinc))*exp(-beta*log(mdatasub$dist)))

```

### 7.1.4 Save the fitted values

#### 7.1.4.1 Run model and save new flow estimates

- Run the model and save all of the new flow estimates in a new column in the dataframe.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$unconstrainedEst2 <- round(mdatasub$unconstrainedEst2,0)
sum(mdatasub$unconstrainedEst2)
```

#### 7.1.4.2 Turn output into matrix

- Use *dcast()* of *maditr** package.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "unconstrainedEst2", margins=c("Orig_code", "Dest_code"))
mdatasubmat2
```

#### 7.1.4.3 Compare with original matrix

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
mdatasubmat
```

#### 7.1.4.4 Visualise actual and estiamted flow

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `unconstrainedEst2`))+
  geom_point(color="black", fill="light blue")
```

### 7.1.5 Assess model performance

- Use *postReSample()* of **caret** package to compute three Goodness-of-Fit statistics To provide a more formal assessment of the model, Goodness-o-Fit statistics 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
postResample(mdatasub$Flow,mdatasub$unconstrainedEst2)

```

Results above show that:

- R-squared value of 0.32 is relatively low.
- Seems that the unconstrained model **failed to fit the empirical data well**.

## 7.2 Origin Constrained Spatial Interaction Model

### 7.2.1 Calibrate model

- Calibrate an origin constrained SIM (the “-1” indicates no intercept in the regression model) by using *glm()*
- *summary()* to look at the summary statistics

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
origSim <- glm(Flow ~ Orig_code+log(wj3_destmedinc)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

summary(origSim)
```

### 7.2.2 Examine how contraints hold for destinations

#### 7.2.2.1 Fit model

- Firstly, we will fitted the model and roundup the estimated values 

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$origSimFitted <- round(fitted(origSim),0)

```

#### 7.2.2.2 Turn paired list into matrix

- To create pivot table to turn paired list into matrix.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat3 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "origSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat3
```

#### 7.2.2.3 Compare with  original observed data

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat
```

#### 7.2.2.4 Display actual and estimated flow

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `origSimFitted`))+
  geom_point(color="black", fill="light blue")
```

### 7.2.3 Assess model

- Compare the fitted values and the actual values by computing Goodness-of-fit statistics.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
postResample(mdatasub$Flow,mdatasub$origSimFitted)

```

Results above show that:

-  R-squared **improved** considerably from 0.32 in the unconstrained model to 0.43 in this origin constrained model.


## 7.3 Destination Constrained Spatial Interaction Model

### 7.3.1 Calibrate model

- Calibrate a destination constrained SIM (the “-1” indicates no intercept in the regression model) by using *glm()*.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
destSim <- glm(Flow ~ Dest_code+log(vi1_origpop)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(destSim)
```

### 7.3.2 Examine how contraints hold for destinations

#### 7.3.2.1 Fit model

- Fit the model and roundup the estimated values

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$destSimFitted <- round(fitted(destSim),0)
```

#### 7.3.2.2 Turn paired list into matrix

- To create pivot table to turn paired list into matrix.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat6 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "destSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat6
```

#### 7.3.2.3 Compare with original observed data

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat
```

#### 7.3.2.4 Display actual and estimated flow

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `destSimFitted`))+
  geom_point(color="black", fill="light blue")
```

### 7.3.3 Assess model

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
postResample(mdatasub$Flow,mdatasub$destSimFitted)

```

Results above show that:

- R-squared **improved further** from 0.43 in the unconstrained model to 0.65 in this origin constrained model

## 7.4 Doubly Constrained Spatial Interaction Model

### 7.4.1 Calibrate model

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
doubSim <- glm(Flow ~ Orig_code+Dest_code+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(doubSim)
```

### 7.4.2 Examine how contraints hold for destinations

#### 7.4.2.1 Fit model

- Fit the model and roundup the estimated values

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasub$doubsimFitted <- round(fitted(doubSim),0)
```

#### 7.4.2.2 Turn paired list into matrix

- To create pivot table to turn paired list into matrix.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat7 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "doubsimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat7
```

#### 7.4.2.3 Compare with original observed data

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
mdatasubmat
```

#### 7.4.2.4 Display actual and estimated flow

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `doubsimFitted`))+
  geom_point(color="black", fill="light blue")
```

Results above show that:

- Fitted values are highly correlated with the actual flow values.
- This show the Doubly Constrained Spatial Interaction Model is the best fit model among the four spatial interaction models.

### 7.4.3 Assess model

- To provide a quantitative assessment of the model, we can compute the Goodness-of-fit statistics exactly the same way as before.

```{r echo=TRUE, eval=TRUE,  cache=TRUE}
postResample(mdatasub$Flow,mdatasub$doubsimFitted)

```

Results above show that:

- The Goodness-of-fit statistics reveal that the Doubly Constrained Spatial Interaction Model is the **best modell** because it produces the **best R-squared statistic** and **smallest RMSE**.


