---
title: "In-class Exercise 5"
description: |
  A short description of the post.
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE, eval=TRUE, echo = TRUE, message = FALSE, error = FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading the R package

- maptools: to create spatial objects
- sf: to import geosptial data
- raster: to create raster object
- spatstat: to perform spatial point pattern analysis (SPPA)
- tmap: to plot thematic maps 
- tidyverse for data handling

```{r eval=TRUE, echo = TRUE}
packages = c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse')

for (p in packages){
if(!require(p, character.only = T)){
  install.packages(p)
}
  library(p,character.only = T)
}
```

## Importing Geospatial Data

- Import shapefile using *st_read()* of **sf** package. Output object is in tibble sf object class.

```{r eval=TRUE, echo = TRUE}
mpsz_sf <- st_read(dsn = 'data/shapefile', layer = 'MP14_SUBZONE_WEB_PL')
  
```

- Projection is in SVY21. 

## Import aspatial data from rds folder

- *read_rds()* of **readr** package is used instead of *readRDS()* of base R is used. This is because output of **read_rds()** is in tibble object. 

```{r eval=TRUE, echo = TRUE}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")

```

- Note that there are some data issue in childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields seem to be in decimal degree. Hence wgs referencing system is assumed.







