---
title: "In-class Exercise 5"
description: |
  In In-class Exercise 5, we revised what was covered in Hands-on Exercise 4, which focuses on spatial point pattern analysis. 
author:
  - name: Wong Wei Ling
    url: www.google.com
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE, eval=TRUE, echo = TRUE, message = FALSE, error = FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading the R package

- maptools: to create spatial objects
- sf: to import geosptial data
- raster: to create raster object
- spatstat: to perform spatial point pattern analysis (SPPA)
- tmap: to plot thematic maps 
- tidyverse for data handling

```{r eval=TRUE, echo = TRUE}
packages = c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse', 'plotly', 'ggthemes')

for (p in packages){
if(!require(p, character.only = T)){
  install.packages(p)
}
  library(p,character.only = T)
}
```

## Importing Geospatial Data

- Import shapefile using *st_read()* of **sf** package. Output object is in tibble sf object class.

```{r eval=TRUE, echo = TRUE}
mpsz_sf <- st_read(dsn = 'data/shapefile', layer = 'MP14_SUBZONE_WEB_PL')
  
```

- Projection is in SVY21. 

## Import aspatial data from rds folder

- *read_rds()* of **readr** package is used instead of *readRDS()* of base R is used. This is because output of **read_rds()** is in tibble object. 

```{r eval=TRUE, echo = TRUE}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")

```

- Note that there are some data issue in childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields seem to be in decimal degree. Hence wgs referencing system is assumed.


## Converting the aspatial data frame into sf objects

```{r eval=TRUE, echo = TRUE}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE", "Y_COORDINATE"),
                    crs = 3414)

```

- Note: *st_as_sf()* accept coordinates in character data type. 

- lat long in decimal degrees

```{r eval=TRUE, echo = TRUE}

childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng", "Lat"),
                    crs = 4326)  %>%
  st_transform(crs = 3414)
```


## Plot to review

- alpha: to set transparency

```{r eval=TRUE, echo = TRUE}
tmap_mode("view")

tm_shape(childcare_sf) +
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha = 0.4,
          col = "red",
          size = 0.05)
```

## Geospatial Data Wrangling

### Convert sf to Spatial* classes

- as_Spatial() of sf package

```{r eval=TRUE, echo = TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)

```

### Convert Spatial* dataframe into Spatial* objects

- as.SpatialPoint() of as.SpatialPolygon() of maptools package

```{r eval=TRUE, echo = TRUE}
childcare_sp <- as(childcare, "SpatialPoints") # data will be dropped. 
CHAS_sp <- as(CHAS, "SpatialPoints") # data will be dropped. 
mpsz_sp <- as(mpsz, "SpatialPolygons") # data will be dropped. 

```

### Convert from Spatial* objects into ppp object

- to drop project info. only X, Y coordinates will be left.
- as.ppp() of maptools package

```{r eval=TRUE, echo = TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")

```

### Remove duplicated points using jitter

- to avoid points overlapping 

```{r eval=TRUE, echo = TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)

any(duplicated(childcare_ppp_jit))
```


```{r eval=TRUE, echo = TRUE}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)

any(duplicated(CHAS_ppp_jit))
```

**Note:**

- must be sf or sp to plot a tmap. 
- For now, we can only plot using sf plot()

### Extract Punggol Planning Area

- Remember to add the comma at the end

```{r eval=TRUE, echo = TRUE}
pg <- mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
```

### Convert SpatialPolygonsDataFrame into SpatialPolygons object

```{r eval=TRUE, echo = TRUE}
pg_sp <- as(pg, "SpatialPolygons")
```

### COnvert SpatialPolygons into owin object

- owin: an object needed by spatstat to define irregular window

```{r eval=TRUE, echo = TRUE}
pg_owin <- as(pg_sp, "owin")
```

### Extract spatial points within owin

```{r eval=TRUE, echo = TRUE}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```


- Plotting with childcare_ppp_jit would not work. Does not show the boundary of Punggol

```{r eval=TRUE, echo = TRUE}
plot(childcare_pg)
```


## L-function 
```{r eval=TRUE, echo = TRUE}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global = TRUE)
```

# Code chunk for plotting interactive L-function

```{r eval=TRUE, echo = TRUE}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```




